---
title: "Insights & Analysis"
output: html_document
---

# Final Project Analysis & Insights

This section provides a deeper statistical and analytical look at U.S. influenza trends from 2000 to 2023, based on the CDC FluView dataset.

---

## Year-over-Year Flu Cases

```{r}
library(tidyverse)
data <- read.csv("cleaned_influenza_data.csv")

yearly_cases <- data %>%
  group_by(Year) %>%
  summarize(Total_Cases = sum(Total_A, na.rm = TRUE) + sum(Total_B, na.rm = TRUE))

ggplot(yearly_cases, aes(x = Year, y = Total_Cases)) +
  geom_line(color = "orange", size = 1) +
  geom_point(color = "red") +
  labs(title = "Total Influenza Cases by Year (2000–2023)",
       x = "Year", y = "Total Cases")
```

---

## Peak Flu Week Analysis

```{r}
peak_weeks <- data %>%
  group_by(Week) %>%
  summarize(Average_Cases = mean(Total_A + Total_B, na.rm = TRUE)) %>%
  arrange(desc(Average_Cases)) %>%
  slice(1:5)

peak_weeks
```

---

## Correlation Between ILI and % Positive Flu Tests

```{r}
cor_test <- cor.test(data$Percent_Positive, data$Weighted_ILI, use = "complete.obs")
cor_test
```

```{r}
ggplot(data, aes(x = Percent_Positive, y = Weighted_ILI)) +
  geom_point(alpha = 0.5, color = "cyan") +
  geom_smooth(method = "lm", color = "yellow") +
  labs(title = "Correlation between % Positive Flu Tests and ILI Percentage",
       x = "% Positive Flu Tests", y = "ILI Percentage")
```

---

## Flu Case Forecasting (ARIMA Model)

```{r}
library(forecast)
library(tibble)

yearly_cases_ts <- ts(yearly_cases$Total_Cases, start = 2000)
model <- auto.arima(yearly_cases_ts)
forecasted <- forecast(model, h = 5)

plot(forecasted, main = "Forecasted Influenza Cases (Next 5 Years)")
```



---

##  Data Overview

The data displayed in this analysis represents weekly influenza surveillance reports collected by the **Centers for Disease Control and Prevention (CDC)** from **2000 to 2023**. It includes laboratory-confirmed influenza cases from clinical and public health labs across the United States, broken down by virus type (Influenza A & B). Additionally, the dataset incorporates information on **Influenza-Like Illness (ILI)**, which measures the percentage of outpatient healthcare visits related to flu symptoms. This combined dataset captures over two decades of flu activity, providing insights into yearly trends, peak flu weeks, and the correlation between confirmed flu cases and ILI reports. The data is summarized and visualized to help users identify long-term patterns, annual fluctuations, and potential forecasting of future influenza seasons.


---

<footer style="font-size: 0.8em; margin-top: 30px;">
Created by Garrett Warr | UVU Data Science Final Project | Spring 2025
</footer>
