---
title: "Misc"
output: html_document
date: "`r Sys.Date()`"
---
 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(palmerpenguins)
library(easystats)
library(equatiomatic)
library(pander)
library(broom)
library(kableExtra)
#install.packages("kableExtra")
#install.packages("pander")
#install.packages("equatiomatic")
 
theme_set(theme_minimal())
 
##command option i short hand for R code##
##look at markdownlivepreview.com to get some help with coding in R markdown
#rmdformats look this up to get templates of websites
 
 
```
 
#Hey look its penguins
 
```{r fig.height=2, fig.width=1, fig.align='center'}

```
 
##Here's a look at the palmer penguins data set:
 
```{r}
penguins %>% glimpse
```
The steps of data science are:
 
1. Import
2. Clean 
3. Visualize 
4. Model 
5. Report
 
We don't need to do any real cleaning for this data set. But we can show some pretty 
pictures...
 
```{r}
p<-
penguins %>%
  filter(!is.na(sex)) %>%
  ggplot(aes(x = species, 
             y = body_mass_g, 
             colour = sex)) + 
  geom_boxplot() + 
  geom_jitter(color='black', alpha=.25, width = .2) + 
  labs(x = "Species", 
       y = "body mass(g)", 
       color = "Sex")
 
plotly::ggplotly(p)
 
```
 
Here, we see that male and female penguins have distinct body mass distributions for each 
species of penguin. Don't believe your eyes?
 
Here are some stats:
 
```{r}
mod1 <- 
penguins %>%
  filter(!is.na(sex)) %>%
  glm(data=.,
      formula = body_mass_g ~ sex * species)
 
```
 
 
```{r}
mod1 %>% tidy %>% 
  kable %>%
  kable_classic(lightable_options = "hover")
```
 
The equation for our model is:
 
```{r}
equatiomatic::extract_eq(mod1)
```