---
title: "Outbreak Insights"
output: 
  html_document:
    theme: journal
    toc: true
    toc-depth: 2
    number-sections: true
---

# Executive Summary

This dashboard analyzes influenza trends in the United States from 2000 to 2023 using CDC FluView surveillance data. Key metrics such as total confirmed cases, influenza-like illness (ILI) rates, and subtype variations are explored. The project also previews flu case forecasts using machine learning (Prophet), providing public health decision-makers with insights on seasonality and strain behavior.

```{r setup, include=FALSE}
library(tidyverse)
library(plotly)
library(ggplot2)
library(RColorBrewer)
library(knitr)

# Load data
data <- read.csv("cleaned_influenza_data.csv")
summary_data <- read.csv("ICL_NREVSS_Summary.csv")
forecast_data <- read.csv("flu_forecast_all_states.csv")

# Safe color palette function
safe_palette <- function(n) {
  RColorBrewer::brewer.pal(max(3, n), "Set2")
}
```

# Methods

The analysis uses weekly surveillance data from the CDC FluView system (2000–2023). ILI metrics represent weighted patient-reported flu-like symptoms. Prophet, an open-source forecasting tool developed by Facebook, was used to predict future flu activity. Time-series smoothing and data cleaning were performed using `tidyverse` packages.

# 1. Total Influenza Cases Over Time

This chart displays the yearly total confirmed cases of Influenza A and B in the United States.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
data %>%
  group_by(Year) %>%
  summarize(
    Total_A = sum(Total_A, na.rm = TRUE),
    Total_B = sum(Total_B, na.rm = TRUE)
  ) %>%
  pivot_longer(cols = c(Total_A, Total_B), names_to = "Type", values_to = "Count") %>%
  plot_ly(x = ~Year, y = ~Count, color = ~Type, type = 'scatter', mode = 'lines+markers') %>%
  layout(title = "Total Influenza A & B Cases by Year",
         xaxis = list(title = "Year"),
         yaxis = list(title = "Confirmed Cases"),
         hovermode = "compare",
         legend = list(orientation = "h", x = 0.1, y = -0.2))
```

**What you're seeing:**  
A comparison of Influenza A and B trends over time. Influenza A typically dominates most seasons, but spikes in B (such as in 2012 and 2018) provide insight into unusual activity.

# 2. ILI (Influenza-Like Illness) Trends

```{r, echo=FALSE, warning=FALSE, message=FALSE}
data %>%
  group_by(Year) %>%
  summarize(Avg_ILI = mean(Weighted_ILI, na.rm = TRUE)) %>%
  plot_ly(x = ~Year, y = ~Avg_ILI, type = 'bar', name = 'Avg ILI') %>%
  layout(title = "Average ILI by Year",
         xaxis = list(title = "Year"),
         yaxis = list(title = "Avg ILI (%)"),
         font = list(family = "Arial", size = 14))
```

**What you're seeing:**  
Weighted ILI reflects flu-like symptoms as reported by patients and providers. It often signals the early rise of flu season before lab-confirmed case surges.

# 3. Top 5 Peak Seasons by Confirmed Cases

```{r, echo=FALSE, warning=FALSE, message=FALSE}
data %>%
  group_by(Year) %>%
  summarize(
    Total_A = sum(Total_A, na.rm = TRUE),
    Total_B = sum(Total_B, na.rm = TRUE),
    Total = Total_A + Total_B
  ) %>%
  arrange(desc(Total)) %>%
  head(5) %>%
  kable(caption = "Top 5 Influenza Seasons by Total Confirmed Cases", digits = 0)
```

**What you're seeing:**  
A table ranking the five flu seasons with the highest number of confirmed cases. Use this for historical comparison and identifying outlier seasons like the 2009 H1N1 pandemic.

# 4. ILI vs. Confirmed Case Correlation

```{r, echo=FALSE, warning=FALSE, message=FALSE}
data %>%
  group_by(Year) %>%
  summarize(
    Total_Cases = sum(Total_A + Total_B, na.rm = TRUE),
    Avg_ILI = mean(Weighted_ILI, na.rm = TRUE)
  ) %>%
  plot_ly(
    x = ~Avg_ILI,
    y = ~Total_Cases,
    type = 'scatter',
    mode = 'markers',
    marker = list(size = 10, color = ~Year, colorscale = 'Viridis'),
    text = ~paste("Year:", Year,
                  "<br>Avg ILI:", round(Avg_ILI, 2),
                  "<br>Total Cases:", Total_Cases),
    hoverinfo = 'text'
  ) %>%
  layout(
    title = "ILI vs. Confirmed Influenza Cases",
    xaxis = list(title = "Average ILI (%)"),
    yaxis = list(title = "Total Confirmed Cases"),
    showlegend = FALSE,
    font = list(family = "Arial", size = 14)
  )
```

**What you're seeing:**  
A scatterplot assessing the relationship between average ILI percentages and total confirmed flu cases by year. Not every ILI spike translates to a high confirmed flu season—but patterns emerge.

# Summary: What the Data Shows

- **Strain shifts matter:** H3N2 and H1N1 alternate, B lineages vary  
- **ILI as an early signal**: Often spikes before confirmed case loads  
- **Seasonal window:** Flu activity consistently peaks in winter months  
- **Forecast use:** Knowing past peaks and subtypes improves model targeting

# References

- Centers for Disease Control and Prevention (CDC). FluView: Weekly U.S. Influenza Surveillance Report. https://www.cdc.gov/flu/weekly/  
- Taylor, R. J., & Prophet Team. (2017). Forecasting at Scale. *arXiv preprint arXiv:1701.07897*.

<footer style="font-size: 0.9em; margin-top: 30px; border-top: 1px solid #ddd; padding-top: 10px; color: #555;">
  <p style="margin: 4px 0;">
    Created by Garrett Warr | <strong>UVU Data Science Capstone</strong> | Spring 2025  
  </p>
  <p style="margin: 4px 0;">
    Data Source: CDC FluView | Educational Use Only  
  </p>
  <p style="margin: 4px 0;">
    Last Updated: April 2025
  </p>
</footer>
